\begin{thebibliography}{99999}

\bibitem[AIIM]{AIIM} Association for Information and Image Management (AIIM).
\emph{What is ECM?}\\
\url{http://www.aiim.org/about-ecm.asp}

\bibitem[ARK03]{ARK03} Atul Ravi Khemuka.
\emph{Workflow Modeling Using Finite Automata}.
PhD Thesis, Department of Industrial and Management Engineering, College of Engineering, University of South Florida, USA, 2003.

\bibitem[BK03]{BK03} Bartosz Kiepuszewski.
\emph{Expressiveness and Suitability of Languages for Control Flow Modelling in Workflows}.
PhD Thesis, Faculty of Information Technology, Queensland University of Technology, Australia, 2003.

\bibitem[DAM01]{DAM01} Dragos-Anton Manolescu.
\emph{Micro-Workflow: A Workflow Architecture Supporting Compositional Object-Oriented Software Development}.
PhD Thesis, Department of Computer Science, University of Illinois at Urbana-Champaign, USA, 2001.

\bibitem[DG95]{DG95} Dimitrios Georgakopoulos and Mark F. Hornick and Amit P. Sheth.
\emph{An Overview of Workflow Management: From Process Modeling to Workflow Automation Infrastructure}.
In: Distributed and Parallel Databases, Volume 3, Number 2, Pages 119--153, 1995.

\bibitem[DQZ01]{DQZ01} Da-Qian Zhang and Kang Zhang and Jiannong Cao.
\emph{A Context-Sensitive Graph Grammar Formalism for the Specification of Visual Languages}.
In: The Computer Journal, Volume 33, Number 3, Pages 186--200, 2001.

\bibitem[DS06]{DS06} Douglas C. Schmidt.
\emph{Model-Driven Engineering}.
In: IEEE Computer, Volume 39, Number 2, Pages 25--31, 2006.

\bibitem[FB96]{FB96} Frank Buschmann and Regine Meunier and Hans Rohnert and Peter Sommerlad and Michael Stahl.
\emph{Pattern-Oriented Software Architecture -- A System of Patterns}.
John Wiley \& Sons, 1996.

\bibitem[FG02]{FG02} Florent Guillaume.
\emph{Trying to unify Entity-based and Activity-based workflows}.\\
\url{http://wiki.zope.org/zope3/TryingToUnifiyWorkflowConcepts}

\bibitem[GF03]{GF03} Garland Foster and Richard Moore and Eduardo Polidor.
\emph{Galaxia: An Open Source Workflow Engine for Tiki}\\
\url{http://workflow.tikiwiki.org/tiki-index.php?page=GalaxiaConcepts}

\bibitem[GoF94]{GoF94} Erich Gamma and Richard Helm and Ralph Johnson and John Vlissides.
\emph{Design Patterns: Elements of Reusable Object-Oriented Software}.
Addison-Wesley, 1994.

\bibitem[JBOSS]{JBOSS} The JBoss Project.
\emph{JBoss jBPM: Workflow and BPM Made Practical}.\\
\url{http://docs.jboss.com/jbpm/v3/userguide/graphorientedprogramming.html}

\bibitem[JD01]{JD01} Jörg Desel and Gabriel Juhás.
\emph{What Is a Petri Net?}.
In: Unifying Petri Nets: Advances in Petri Nets, Lecture Notes in Computer Science,
Volume 2128/2001, Springer, 2001.

\bibitem[JM01]{JM01} Jishnu Mukerji and Joaquin Miller.
\emph{MDA Guide}.\\
\url{http://www.omg.org/cgi-bin/apps/doc?omg/03-06-01.pdf}

\bibitem[KL97]{KL97} Karl J. Lieberherr.
\emph{Connections between Demeter/Adaptive Programming and Aspect-Oriented Programming (AOP)}.
College of Computer Science, Northeastern University, Boston, MA, USA, 1997.\\
\url{http://www.ccs.neu.edu/home/lieber/connection-to-aop.html}

\bibitem[MF05]{MF05} Martin Fowler.
\emph{Language Workbenches: The Killer-App for Domain Specific Languages?}
June, 2005.\\
\url{http://martinfowler.com/articles/languageWorkbench.html}

\bibitem[ML05]{ML05} Markus Löschnigg.
\emph{XML in Workflow Management Systems}.
Master's Thesis, Graz University of Technology, Austria, 2005.

\bibitem[OPENFLOW]{OPENFLOW} The OpenFlow Project.
\emph{OpenFlow: An Introduction}.\\
\url{http://www.openflow.it/wwwopenflow/Documentation/documentation/OpenFlowIntroduction/}

\bibitem[PM99]{PM99} Peter Muth and Jeanine Weisenfels and Michael Gillmann and Gerhard Weikum.
\emph{Integrating Light-Weight Workflow Management Systems within Existing Business Environments}.
In: Proceedings of the 15th International Conference on Data Engineering, March 1999, Sydney, Australia.

\bibitem[RA01]{RA01} Rob Allen.
\emph{Workflow: An Introduction}.
In: Workflow Handbook, Workflow Management Coalition, 2001.

\bibitem[RF00]{RF00} Robert E. Filman and Daniel P. Friedman.
\emph{Aspect-Oriented Programming is Quantification and Obliviousness}.
In: Proceedings of the Workshop on Advanced Separation of Concerns, OOPSLA 2000, October 2000, Minneapolis, USA.

\bibitem[SB05]{SB05} Sebastian Bergmann.
\emph{Professionelle Softwareentwicklung mit PHP 5}.
dpunkt.verlag, 2005.

\bibitem[SB06]{SB06} Sebastian Bergmann and Günter Kniesel.
\emph{GAP: Generic Aspects for PHP}.
Third European Workshop on Aspects in Software, August 2006, University of Twente, Enschede, The Netherlands.

\bibitem[SF04]{SF04} Sérgio Fernandes and João Cachopo and António Rito-Silva.
\emph{Supporting Evolution in Workflow Definition Languages}.
In: Proceedings of the 20th Conference on Current Trends in Theory and Practice of Computer Science (SOFSEM 2004), Springer-Verlag, 2004.

\bibitem[SH05]{SH05} Saphira Heijens.
\emph{Support for Workflow Administration and Monitoring in the YAWL Environment}.
Master Thesis, Vrije Universiteit, Amsterdam, The Netherlands, 2005.

\bibitem[TM04]{TM04} Tony Marston.
\emph{An Activity-Based Workflow Engine for PHP}.
September 2004.\\
\url{http://www.tonymarston.net/php-mysql/workflow.html}

\bibitem[TS04]{TS04} Thomas Schmidt.
\emph{Erweiterung und Integration des Open Source Workflow-Systems SWAMP am Beispiel eines Software-Wartungs-Prozesses}.
Diploma Thesis, Georg-Simon-Ohm-Fachhochschule, Nürnberg, Germany, 2004.

\bibitem[W3C07]{W3C07} World Wide Web Consortium.
\emph{XSL Transformations (XSLT) Version 2.0}.
W3C Recommendation, January 2007.\\
\url{http://www.w3.org/TR/2007/REC-xslt20-20070123/}

\bibitem[WA96]{WA96} W. M. P. van der Aalst.
\emph{Petri-net-based Workflow Management Software}.
In: Proceedings of the NFS Workshop on Workflow and Process Automation in Information Systems, May 1996, Athens, Georgia, USA.

\bibitem[WA04]{WA04} W. M. P. van der Aalst and L. Aldred and M. Dumas and A. H. M. ter Hofstede.
\emph{Design and Implementation of the YAWL System}.
In: Proceedings of the 16th International Conference on Advanced Information Systems Engineering (CAiSE 2004), June 2004, Riga, Latvia.

\bibitem[WfMC95]{WfMC95} Workflow Management Coalition.
\emph{The Workflow Reference Model}.
Document Number WFMC-TC-1003, 1995.

\bibitem[WfMC99]{WfMC99} Workflow Management Coalition.
\emph{Terminology and Glossary}.
Document Number WFMC-TC-1011, 1999.

\bibitem[WfMC05]{WfMC05} Workflow Management Coalition.
\emph{Workflow Process Definition Interface -- XML Process Definition Language (XPDL)}.
Document Number WFMC-TC-1025, 2005.

\end{thebibliography}
